Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-17 Cost_manuscript.Rmd
working a bit on discussion.

  What is often missing from existing accounts is the common case of choosing among types of behavioral demands. Studies have traditionally compared the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [@Ainslie1975; @Frederick2002; @Green1994; @Kable2007; @Kool2010; @Westbrook2013]. Although this line of work suggests that discounting functions are similar for delay and effort [@Massar2015; @Prevost2010; @Seaman2018], there is ongoing debate about the shape of the discounting function across effort types [@Arulpragasam2018; @Chong2017; @Klein-Flugge2016; @Kool2018; @Prevost2010], and discount rates do not always correlate across cost domains [@Seaman2018; @Westbrook2013]. However, sporadic preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context [@Kool2018]. For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting (e.g. taking the cumbersome way home rather than waiting for a delayed train, even if the resulting time of arrival is the same). Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available [@Olivola2013]. Recent work has proposed that the subjective costs produced by delays and cognitive effort is modulated by the possibility of missing out on better concurrent alternatives, namely the opportunity cost of time [OC; @Fawcett2012; @Kurzban2013; @Otto2019]. This evidence sets the stage for examining what people would actually select when given the option for multiple types of demands, including the unexplored question of deciding between cognitive and physical effort [@Inzlicht2018; @Kool2018].

@@

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could depend on the ability to achieve the same outcome through different means (**not quite**). Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

  Our findings provided partial support for our pre-registered hypotheses, which can be found in https://osf.io/2rsgm/. In accord with our predictions, decision makers were successful in approximating reward-maximizing strategies, as evidenced by acceptance rates. This is in line with previous foraging experiments, where the main deviation from optimality is a tendency to overharvest. Unlike our first hypothesis, our predictions regarding differences among costs did not

  Confirming our predictions, we found that individuals could approximate optimality by preferring higher rewards and shorter delays, regardless of the richness of the environment or whether they faced a single or intermixed demands. When faced with a single type of demand, people in the cognitive effort group accepted more trials than optimal, whereas physical effort and wait groups under-accepted, partially matching our predictions and highlighting that cognitive effort can add value to outcomes. However, against our expectations, we found that a similar preference for cognitive effort disappeared over time as decision makers foraged for types of demands in addition to rewards. Computational modeling suggested that cognitive effort was valuable because it shortened the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands. The shorter time sensitivity of cognitive effort could be due to its constant recruitment of working memory, impeding estimations of the nominal time, or because of the discrete nature of the task (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios.

>>>>

  What is often missing from existing accounts is the common case of choosing among types of behavioral demands. Studies have traditionally compared the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [@Ainslie1975; @Frederick2002; @Green1994; @Kable2007; @Kool2010; @Westbrook2013]. Although this line of work suggests that discounting functions are similar for delay and effort [@Massar2015; @Prevost2010; @Seaman2018], there is ongoing debate about the shape of the discounting function across effort types [@Arulpragasam2018; @Chong2017; @Hartmann2013; @Klein-Flugge2016; @Kool2018; @Prevost2010], and discount rates do not always correlate across cost domains [@Seaman2018; @Westbrook2013]. However, sporadic preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context [@Kool2018]. For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting (e.g. taking the cumbersome way home rather than waiting for a delayed train, even if the resulting time of arrival is the same). Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available [@Olivola2013]. Recent work has proposed that the subjective costs produced by delays and cognitive effort is modulated by the possibility of missing out on better concurrent alternatives, namely the opportunity cost of time [OC; @Fawcett2012; @Kurzban2013; @Otto2019]. This evidence sets the stage for examining what people would actually select when given the option for multiple types of demands, including the unexplored question of deciding between cognitive and physical effort [@Inzlicht2018; @Kool2018].

@@

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

  Our findings provided partial support for our pre-registered hypotheses, which can be found in https://osf.io/2rsgm/registrations. In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal points. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [constantino; hayden; wikenheiser; sweiss]. This successful integration of the task was further evidenced by the ability of participants to post-experimentally describe near-optimal strategies, discarding the possibility of our results being driven by noise or bias in the estimation of environmental richness [hayden; garrett; dundon]. In contrast, our prediction that effort requirements would produce more acceptances was only true for the cognitive group, and this tendency for higher acceptances did not switch to passive delay as predicted for the within-subjects experiment. These hypotheses originally stemmed from a combination of pilot data, and prevalent theorizing that incentivized work should become more appealing [Eisenberger; Inzlicht], but that people would opt for the easiest available way to achieve desired outcomes [@Hull1943; Olivola]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted task-specific attributes contributing to the attractive or aversive nature of demands. The fact that preferences for cognitive effort were stable as predicted in the first experiment, and present at the beginning of the second experiment, further supported this idea. However, the gradual convergence in acceptances in the second experiment emphasized that these features could be contextually molded.

  These choice dynamics ruled out a number of explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. Effort preferences due to boredom would not have shown the global acceptance decline and convergence in the within-subjects experiment, which also discards the possibility that all our participants were high on need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Even though the reciprocal relationship between effort and outcome can often explain preferences for effort (learned industriousness, Inzlicht2018), the observed differences in acceptances between effort types and the eventual convergence of rates in the within-subjects study cannot be explained by reward alone, as demands were rewarded at the same rate.

  We do not discard the influence of these elements in daily effort evaluations. Instead, we computationally highlight that effort can became valuable by shortening the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands. Alterations in subjective time sensitivity have been shown to play a role in temporal discounting [@Mckerchar2003], and are often experienced when we are fully immersed in a task [@czksdmsdkjdc]. The shorter time sensitivity of cognitive effort could be due to its constant recruitment of working memory, impeding estimations of the nominal time, or because of the discrete nature of the task (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios.

  The use of foraging paradigms....
